### 2012-08-01 trunk/ijcis-pos-model/sources/temporalModel.tex
preprint version

In this section we will formalize the model for possibilistic valid-time relational databases. The first subsection is devoted to the formalization of the model. Then, the data manipulation language is defined.

\subsection{\label{subsec:temporal-model}The generalized temporal model}
The model is based on GEFRED~\cite{Medina1994} (Generalized Model of Fuzzy Relational DB) model. This model is extended by adding valid-time support which will be illustrated through definitions and examples. The information in the system is defined by the following elements:

@@

\begin{definition}
\label{def:generalized-fuzzy-temporal-domain}
\emph{Generalized fuzzy temporal domain.}
Consider $\T$ to be the temporal domain, and let $\tilde \Pow\left( \T\right)$ be the set of all \emph{normalized} possibility distributions (see Section \ref{subsec:possibility-theory}, equation \eqref{NormalizationProperty}) defined on $\T$.
The Generalized Fuzzy Temporal Domain, $\T_G$ is
\begin{equation}
\T_G \subseteq \left \lbrace \tilde \Pow\left( \T\right) \cup \text{NULL} \right \rbrace
\end{equation}
\end{definition}

@@

\item We will note by $r$ any valid instance of $R_{FTG}$. 
      \begin{equation}
       \label{eq:fuzzy-valid-time-instance}
      r \subseteq D_1\ x\ \ldots\ x\ D_n
      \end{equation}

@@

We will illustrate the definitions with an example.

\begin{example}
Consider the set of attributes $A = \left \lbrace A_1, A_2, A_3 \right \rbrace$. The primary key for these attributes is given by $A_K = \left \lbrace A_1, A_2 \right \rbrace$. Let $I = \left \lbrace S, E \right \rbrace$ be the set of temporal attributes that define the possibilistic validity period of the data. $R_{FTG} = A \cup V_{ID} \cup I$ is the valid-time relation and $r$ is an instance of the relation. The instance $r$ is given by the following elements. $r = \left \lbrace \left(a_{11}, a_{12}, a_{13}, 001, s_1 ,e_1 \right) \right.$,  $\left(a_{21}, a_{22}, a_{23}, 001, s_2, e_2 \right)$ , $\left. \left(a_{11}, a_{12}, a_{31}, 002, s_3, e_3 \right) \right \rbrace$. The instance $r$ is illustrated in Table \ref{tbl:fuzzy-sample-definitions}. 
Consider the tuple $t = \left(a_{11}, a_{12}, a_{13}, s_1 ,e_1 \right)$. Then,
\begin{align}
 \nonumber
s_1 &= t\left[S \right]\\
 \nonumber
e_1 &= t[E]\\
 \nonumber
i\ \ &= \left(s_1, e_1\right)\\
 \nonumber
a_k &= t\left[A_K\right] = \left(a_{11}, a_{12} \right) \\
 \nonumber
k &= t\left[PK\right] =\left(a_{11}, a_{12}, s_1, e_1\right)\\
 \nonumber
Vt &= r\left(t\left[A_K\right] \right) = \left \lbrace t_1, t_3 \right \rbrace
\end{align}

@@

% 
% 
% Consider $\A$ to be the set of all the entities, and let $A=\left(A_1, \ldots, A_n \right), A \in \A$ be the set of (fuzzy) attributes that define an entity. Let $\I_{PVP}$ be the set of all the ill-known time intervals and let $I = \left(X, Y \right)$ be a PVP, $I \in \I_{PVP}$.  The pair $\left(A, I\right)$ expresses that the data regarding the entity A are valid during the ill-known time interval I. Let $R_{FTG} \subseteq \A \  x\  \I_{PVP}$ be a valid-time relation. Then, the following equation indicates that the pair $\left(A, I\right)$ is in the relation R:
% 
% \begin{equation}
% \label{eq:fuzzy-rel-def}
% \left( R_{FTG}, \left(A , I\right) \right)
% \end{equation}
% 
 A generalized fuzzy temporal relation $R_{FTG}$ can be noted also by:
\label{def:generalized-fuzzy-temporal-relation}
\begin{equation}
\label{eq:generalized-fuzzy-temporal-relation}
R_{FTG} = \left(\Head, \Body \right)
\end{equation}
Where $\Head$ is the Head of the relation and consist on a fixed set of triplets attribute- domain - compatibility with an optional the valid-time attribute:

@@

$\Body$ is the body of the relation and it consists on a set of tuples. Each tuple is a triplet attribute- value- degree with an optional valid-time attribute:

\begin{align}
\label{eq:body-valid-time}
\Body = \big \lbrace \left(A_{G1}:\tilde{d}_{i1}\left[,c_{i1} \right] \right),\\
\nonumber
 \ldots,\\
 \nonumber
  \left(A_{Gn}:\tilde{d}_{in}\left[,c_{in} \right] \right),\\
  \nonumber
   \Big[  \left( \text{PVP}, \tilde{d}_{\text{PVP}} \left[,C_{A_{\text{PVP}}} \right] \right)  \Big] \big \rbrace
\end{align}

@@

\begin{definition}
\label{def:value-component}
\emph{Value component.}
The value component $R^{v}_{FTG}$ of a fuzzy temporal relation $R_{FTG}$ is a set with the value components for both the head and the body of the relation:
\begin{align}
\label{eq:value-component}
R^{v}_{FTG} = \left \lbrace \Head^{v},\Body^{v} \right \rbrace \\
\nonumber
\text{Where: } \\
\nonumber
\Head^{v} = \left \lbrace A_{G1}:D_{G1}, \ldots,  A_{Gn}:D_{Gn} \right \rbrace \\
\nonumber
\Body^ {v} = \left \lbrace A_{G1}:\tilde{d}_{i1}, \ldots,  A_{Gn}:\tilde{d}_{in}\ \right \rbrace 
\end{align}
\end{definition}

For example, in the case of the document with ID = 3:
\begin{align}
%\label{eq:value-component-s}
\nonumber
\Head^{v} &=  \lbrace \left(\text{ID}:D_{\text{ID}}\right), \left(\text{Entity}:D_{\text{entity}}\right),\\
\nonumber
& \left(\text{PVP}:D_{\text{PVP}}\right) \rbrace \\
\nonumber
\Body^ {v} &= \left \lbrace 3:, \text{``E.U.''}: ,  \left[15/3/1012, 30/3/1012\right] \right \rbrace 
\end{align}

@@

For example, in the case of the document with ID = 3:
\begin{align}
%\label{eq:sample-temporal-component}
\nonumber
\Head^{t} = & \lbrace \left( \text{PVP}, D_{\text{PVP}}\right),\\
\nonumber
\Body^ {t} =& \left \lbrace 4/4/1012 , UC \right \rbrace 
\end{align}

@@

\begin{definition}
\label{def:generalized-fuzzy-temporal-key}
\emph{Generalized fuzzy temporal key.}
Consider $D_G$ to be a fuzzy generalized domain, and let $A_{Gs}:D_{Gs}$ be the attributes and the domain of the attribute for each $s \in S \subseteq \left(1, \ldots, n \right)$. Let  $V$ be a new attribute called \emph{version}. 
A generalized fuzzy temporal key, $K_{GT}$ is a subset of the head.
\begin{align}
\label{eq:generalized-fuzzy-temporal-key}
K_{GT} \subseteq \Head, K_{GT} = \left \lbrace  \left(A_{Gs}:D_{Gs} \right) \right.  \\
\nonumber
 \left. \cup  \left(V_{ID}:D_{ID} \right)	\right \rbrace \\
\nonumber
\text{Typeof }\left(D_{ID} \right) = \\
\nonumber
s\in S \subseteq \left(1, \ldots, n \right) \\
\end{align}
Subject to the following constraints:
\begin{align}
\forall s \in S, \text{Typeof } \left(D_{Gs} \right) \in \left \lbrace 1, 2 \right \rbrace \\
\nonumber
\forall i, i' \in \left \lbrace 1, \ldots, m\right \rbrace , \exists s \in S: \\
\nonumber
\left(A_{Gs}:d_{is} \right) \neq \left(A_{Gs}:d_{i's} \right)
\end{align}
\end{definition}

@@

The Generalized Fuzzy Relational Algebra~\cite{Medina1994} manipulates relations like $R_{FTG}$. The operations defined are: \emph{Union, Intersection, Difference, Cartesian Product, Projection, Join} and \emph{Selection}. Thus, in this section we will describe and implement the following operations for temporal databases, as described in \ref{subsubsec:Understanding-valid-time-databases}. The operations implemented are: \emph{Insert, Modify} and \emph{Delete}. The semantics of the operations will be the same that those defined for a crisp temporal database, whereas the temporal representation is made by the possibilistic valid-time period and the ill-known constraints (see sections \ref{sec:prelim} and \ref{sec:time-rep}.
%The algebra for the data manipulation is defined in .

It is important to notice that while the result of the evaluation of any comparison between crisp time intervals is boolean, the evaluation of any comparison between \emph{PVPs} is a value in the unit interval. Therefore, for a crisp valid-time relation, it is not possible an overlapping among any valid-time interval. In other words, a crisp temporal database is always strictly consistent. Thus, from the point of view of a fuzzy temporal database, there are the following choices (see Figure \ref{fig:tdb-class}):

% \begin{itemize}
% \item 
\begin{definition}
\textbf{Strictly consistent} It is not possible that two versions of an entity to be valid at the same time.
Consider a tuple $t$ in the instance $r$ of the fuzzy valid-time relation $R_{FTG}$ and  $Vt = \left \lbrace t_1, \ldots, t_n \right \rbrace$ as defined in \ref{def:fuzzy-temporal-relation}. Then, a fuzzy temporal database is strictly consistent if: 
\begin{align}
\label{eq:stricly-consistent}
\forall k,j \in \left \lbrace 1, \ldots, n \right \rbrace, k < j\\
\nonumber t_k \mbox{ Overlaps } t_j \leq 1
\end{align}
\end{definition}

%\item 
\begin{definition}
 \textbf{Ill-consistent} The overlapping of two PVPs of the same entity do overlap. It can be distinguished among the following sub-types:
\end{definition}
% 	\begin{itemize}
% 	\item 
	      \begin{definition}
	      \textbf{Co-existence} Two versions of the same entity may exist at the same time.
		      \begin{align}
		      \forall k, \exists !j \in \left \lbrace 1, \ldots, n \right \rbrace, k < j\\
		      \nonumber t_k \mbox{ Overlaps } t_j \geq 1
		      \end{align}
	      \end{definition}

	%\item 
	      \begin{definition}
	      \textbf{Weak-consistence} Several versions of the same entity may exist at the same time.
		      \begin{align}
		      \label{eq:weak-consistence}
		      \forall k, \exists j \in \left \lbrace 1, \ldots, n \right \rbrace, k < j\\
		      \nonumber t_k \mbox{ Overlaps } t_j \geq 1
		      \end{align}
	      \end{definition}

	%\end{itemize}

\begin{samepage}
\vspace*{13pt}
\begin{center}
{
\includegraphics[scale=1.25]{./graphs/tdb-clas.pdf}

}
\end{center}
%\centerline{ \psfig{file=./graphs/Y-time-point.eps}}
\vspace*{10pt}
\fcaption{\label{fig:tdb-class}Classification for a fuzzy temporal database. The validity for an entity is represented by a trapezoid. Graphic (a) shows an example of strictly consistent database. Between two versions of the entity it is possible some amount of overlapping, but always holds equation \eqref{eq:stricly-consistent}. Graphic (b) shows an example of a co-existence. Two versions of the same entity may overlap in any degree, but not more than two versions. Graphic (c) shows an example of weak-consistence. The different versions of the same entity are shown in an unfolded way, to clarify the multiple overlapping.   }
\vspace*{13pt}
\end{samepage}

%\end{itemize}

In the implementation of the DML operations we will consider a strictly consistent temporal database. This is a natural extension of the crisp temporal databases. This approach allows an implementation for the DML operations defined in Section \ref{subsubsec:Understanding-valid-time-databases}.

@@

\begin{definition}
\label{def:close-r-a-pvp}
\textbf{CloseR$\left(i_1, i_2\right)$.}
Consider the elements in definition \ref{def:generalized-fuzzy-temporal-relation}. The CloseR function closes the PVP given by $i_1$ with a conjunctive combination of ill-known constraints (see section \ref{subsec:interval-evaluation-by-ill-known-constraints}):
\begin{align}
\label{eq:close-r-a-pvp}
\text{CloseR}\left(i_1, i_2\right) &= \\
\begin{cases}
\nonumber
i_1 = \left(s_1, e_z \right) & \mbox{ if } i_1 = (s_1, UC), \\
& e_z \triangleq \left \lbrace C_1\left(>, s_1 \right), C_2\left(<, s_2 \right) \right \rbrace \\
i_1 & \mbox{ otherwise }
% \nonumber
% I   & \mbox{ if } I \neq \left[X, UC \right] \\
% I = \left[X, Z \right] & \mbox{ if } I = \left[X, UC \right] \\
% & Z \triangleq \left \lbrace C_1\left(>, X \right), C_2\left(<, X' \right) \right \rbrace
\end{cases}
\end{align}
\end{definition}

For example, consider $i_1 = \left[\left[4/4/1012, 3, 3\right] , UC\right]$ and $i_2 = \left[ \left[15/4/1012,2,1\right], UC\right]$. The result for CloseR$\left(i_1, i_2\right)$  is $i_1 = \left[ \left[4/4/1012, 3, 3\right], \left[4,7,13,15 \right] \right]$. The value that closes $i_1$ is a trapezoid in the form $\left[ \alpha, \beta, \gamma, \delta \right]$ as defined in Section \ref{subsec:fuzzy-numbers}.

@@

\begin{eqnarray}
\label{eq:pvp-close-current}
\text{Close-current} \left(r, t \right) =\\
\begin{cases}
\nonumber
r - t_{CUR} \cup \lbrace t_{UP}\rbrace \cup \lbrace t \rbrace
& \mbox{ if } t_{CUR} \neq \emptyset \\
% R - \left(A, I \right) \cup \left \lbrace \left(A, \mbox{ CloseR } \left(I, J\right) \right)\right \rbrace \cup \left \lbrace\left(A, J\right) \right \rbrace  \\
% \nonumber
% \mbox{ if  Current } \left(R, A \right) \neq \emptyset\\ %\wedge J > \mbox{ CloseR } \left(I, J \right)   \\
\nonumber r , \text{ otherwise}
\end{cases}
\end{eqnarray}
\end{definition}

@@

\item The entity is in the relation. Depending on the value of the time interval $i$, there are three possibilities:
	\begin{enumerate}
	\item Insert $t$ in the instance $r$ of the relation $R_{FTG}$. If the time interval $i$ does not overlap any other valid-time interval in the relation $R_{FTG}$. Note that here, the result of the Overlaps operator is in the unit interval.
For example, the document with ID=3 is still valid. The insert sentence is the following.

@@

	\item Reject the insertion, if the time interval $i$ do overlap with a degree of 1 any existing valid-time interval for the entity in the relation. For example, consider now that the document manager wants to introduce a past validity period for the document with ID=3. The validity starting date for the document was around 6/4/12/1012 and the ending date was around 25/4/1012. As this interval does overlaps other time intervals with a degree of 1, it is not possible that the document was valid during two different periods of time. Therefore, the insertion is rejected. The insert sentence is:

@@

\begin{align}
\label{eq:insert-fuzzy-temporal}
\INS \left(r, t\right) &=\\
\begin{cases}
\nonumber
r \cup \lbrace t \rbrace
& \mbox{ if }  t \not \in r \mbox{ or }\\
\nonumber & \forall t_k \in Vt,  \left(i \mbox{ Overlaps } i_k\right) < 1\\
% 
% R \cup \left(A, I \right) & \mbox{ if }  (A, I ) \not \in R  \mbox{ or } \\
%  &   \left(A, J\right) \in R \mbox{ and } \lnot \left(I \mbox{ Overlaps } J\right) \\
\mbox{modify(r, t)} & \mbox{ otherwise }    %\mbox{ if } \exists J \left(A,J\right) \in R \wedge \mbox{Current(A,J)} \\ 
%R, &   
\end{cases} 	
\end{align}
\subsubsection{\label{subsubsec:delete-fuzzy-temporal}Delete}
The delete operation logically removes an entity which is valid in the instance $r$ of the relation $R_{FTG}$.

@@

\begin{align}
\label{eq:delete-fuzzy-temporal}
\DEL \left(r, t\right) =
\begin{cases}
\nonumber
r - Vt
\end{cases} 	
\end{align}

>>>>

In this section we will formalize the model for possibilistic valid-time relational databases. The first subsection is devoted to the formalization of the model. Then, a data manipulation language is defined.

\subsection{\label{subsec:temporal-model}The generalized temporal model}
The model is based on GEFRED~\cite{Medina1994} (Generalized Model of Fuzzy Relational DB) model. This model is extended by adding valid-time support which will be illustrated through the following definitions and examples. The information in the system is defined by the following elements:

@@

\begin{definition}
\label{def:generalized-fuzzy-temporal-domain}
\emph{Generalized fuzzy temporal domain.}
Consider $\T$ to be the temporal domain, and let $\tilde \Pow\left( \T\right)$ be the set of all \emph{normalized} possibility distributions defined on $\T$.
The Generalized Fuzzy Temporal Domain, $\T_G$ is
\begin{equation}
\T_G \subseteq \left \lbrace \tilde \Pow\left( \T\right) \cup \text{NULL} \right \rbrace
\end{equation}
\end{definition}

@@

\item We will note by $r$ any valid instance of $R_{FTG}$. 
      \begin{equation}
       \label{eq:fuzzy-valid-time-instance}
      r \subseteq D_1\ x\ \ldots\ x\ D_n x\ \mathbb{N}  x\  \T_G\  x\ \T_G 
      \end{equation}

@@

Table \ref{tbl:fuzzy-sample-definitions} contains an example instance.

@@

% 
% 
% Consider $\A$ to be the set of all the entities, and let $A=\left(A_1, \ldots, A_n \right), A \in \A$ be the set of (fuzzy) attributes that define an entity. Let $\I_{PVP}$ be the set of all the ill-known time intervals and let $I = \left(X, Y \right)$ be a PVP, $I \in \I_{PVP}$.  The pair $\left(A, I\right)$ expresses that the data regarding the entity A are valid during the ill-known time interval I. Let $R_{FTG} \subseteq \A \  x\  \I_{PVP}$ be a valid-time relation. Then, the following equation indicates that the pair $\left(A, I\right)$ is in the relation R:
% 
% \begin{equation}
% \label{eq:fuzzy-rel-def}
% \left( R_{FTG}, \left(A , I\right) \right)
% \end{equation}
% 
 A generalized fuzzy temporal relation $R_{FTG}$ can be noted also by:
\label{def:generalized-fuzzy-temporal-relation}
\begin{equation}
\label{eq:generalized-fuzzy-temporal-relation}
R_{FTG} = \left(\Head, \Body \right)
\end{equation}
Where $\Head$ is the Head of the relation and consists on a fixed set of triplets attribute- domain - compatibility with an optional the valid-time attribute:

@@

$\Body$ is the body of the relation and it consists on a set of tuples. Each tuple is a set of triplets attribute- value- degree with an optional valid-time attribute:

\begin{align}
\label{eq:body-valid-time}
\Body = \big \lbrace \ldots \{ \left(A_{G1}:\tilde{d}_{i1}\left[,c_{i1} \right] \right),\\
\nonumber
 \ldots,\\
 \nonumber
  \left(A_{Gn}:\tilde{d}_{in}\left[,c_{in} \right] \right),\\
  \nonumber
   \Big[  \left( \text{PVP}, \tilde{d}_{\text{PVP}} \left[,C_{A_{\text{PVP}}} \right] \right)  \Big] \} \ldots \big \rbrace
\end{align}

@@

\begin{definition}
\label{def:value-component}
\emph{Value component.}
The value component $R^{v}_{FTG}$ of a fuzzy temporal relation $R_{FTG}$ is a set with the value components for both the head and the body of the relation:
\begin{align}
\label{eq:value-component}
R^{v}_{FTG} = \left \lbrace \Head^{v},\Body^{v} \right \rbrace \\
\nonumber
\text{Where: } \\
\nonumber
\Head^{v} = \left \lbrace (A_{G1}:D_{G1}), \ldots,  (A_{Gn}:D_{Gn}) \right \rbrace \\
\nonumber
\Body^ {v} = \left \lbrace (A_{G1}:\tilde{d}_{i1}), \ldots,  (A_{Gn}:\tilde{d}_{in})\ \right \rbrace 
\end{align}
\end{definition}

For example, in the case of the document with ID = 3:
\begin{align}
%\label{eq:value-component-s}
\nonumber
\Head^{v} &=  \lbrace \left(\text{ID}:D_{\text{ID}}\right), \left(\text{Entity}:D_{\text{entity}}\right),\\
\nonumber
& \left(\text{PVP}:D_{\text{PVP}}\right) \rbrace \\
\nonumber
\Body^ {v} &= \{\ldots \{ (\text{ID}:3), (\text{Entity}:\text{``E.U.''}) ,\\
\nonumber
& t(\text{PVP}:\left[15/3/1012,5,2\right], \left[30/3/1012,1,1\right]) \} \ldots\}
\end{align}

@@

For example, in the case of the document with ID = 3:
\begin{align}
%\label{eq:sample-temporal-component}
\nonumber
\Head^{t} = & \lbrace \left( \text{PVP}, D_{\text{PVP}}\right)\}\\
\nonumber
\Body^ {t} =& \{ \ldots (\text{PVP}:[\left[15/3/1012,5,2\right],\\
\nonumber
& \left[30/3/1012,1,1\right]]) \ldots\} 
\end{align}

@@

\begin{definition}
\label{def:generalized-fuzzy-temporal-key}
\emph{Generalized fuzzy temporal key.}
Consider $D_G$ to be a fuzzy generalized domain, and let $A_{Gs}:D_{Gs}$ be the attributes and the domain of the attribute for each $s \in S \subseteq \left(1, \ldots, n \right)$. Let  $V$ be a new attribute called \emph{version}. 
A generalized fuzzy temporal key, $K_{GT}$ is a subset of the head.

\begin{align}
\label{eq:generalized-fuzzy-temporal-key}
K_{GT} \subseteq \Head, K_{GT} = \left \lbrace \left(A_{Gs}:D_{Gs} \right)  \right \rbrace \\
\nonumber
\cup  \{\left(V_{ID}:\mathbb{N}  \right)\}\\
\nonumber
s\in S \subseteq \left(1, \ldots, n \right)
\end{align}
Subject to the following constraints:
\begin{align}
\forall s \in S, \text{Typeof } \left(D_{Gs} \right) \in \left \lbrace 1, 2 \right \rbrace \\
\nonumber
\forall i, i' \in \left \lbrace 1, \ldots, m \right \rbrace , \exists s \in S: \\
\nonumber
\left(A_{Gs}:d_{is} \right) \neq \left(A_{Gs}:d_{i's} \right)
\end{align}
\end{definition}

@@

The Generalized Fuzzy Relational Algebra~\cite{Medina1994} manipulates relations like $R_{FTG}$. The operations defined are: \emph{Union, Intersection, Difference, Cartesian Product, Projection, Join} and \emph{Selection}. Thus, in this section we will describe and implement the following operations for temporal databases, as described in \ref{subsubsec:Understanding-valid-time-databases}. The operations implemented are: \emph{Insert, Modify} and \emph{Delete}. The semantics of the operations will be the same that those defined for a crisp temporal database, whereas the temporal representation is made by the possibilistic valid-time period and the ill-known constraints (see sections \ref{sec:prelim} and \ref{sec:time-rep}).
%The algebra for the data manipulation is defined in .

It is important to notice that while the result of the evaluation of any comparison between crisp time intervals is boolean, the evaluation of any comparison between \emph{PVPs} is a value in the unit interval.

@@

\begin{definition}
\label{def:close-r-a-pvp}
\textbf{CloseR$\left(i_1, i_2\right)$.}
Consider the elements in definition \ref{def:generalized-fuzzy-temporal-relation} and ill-known intervals given by $i_1=(s_1,e_1)$ and $i_2=(s_2,e2)$. The CloseR function closes the PVP given by $i_1$ with a conjunctive combination of ill-known constraints (see section \ref{subsec:interval-evaluation-by-ill-known-constraints}):
\begin{align}
\label{eq:close-r-a-pvp}
\text{CloseR}\left(i_1, i_2\right) &= \\
\begin{cases}
\nonumber
\left(s_1, e_z \right) & \mbox{ if } i_1 = (s_1, UC), \\
& e_z \triangleq \left \lbrace C_1\left(>, s_1 \right), C_2\left(<, s_2 \right) \right \rbrace \\
i_1 & \mbox{ otherwise }
% \nonumber
% I   & \mbox{ if } I \neq \left[X, UC \right] \\
% I = \left[X, Z \right] & \mbox{ if } I = \left[X, UC \right] \\
% & Z \triangleq \left \lbrace C_1\left(>, X \right), C_2\left(<, X' \right) \right \rbrace
\end{cases}
\end{align}
\end{definition}

For example, consider $i_1 = \left[\left[4/4/1012, 3, 3\right] , UC\right]$ and $i_2 = \left[ \left[15/4/1012,2,1\right], UC\right]$. The result for CloseR$\left(i_1, i_2\right)$  is $i_1 = \left[ \left[4/4/1012, 3, 3\right], \left[4,7,13,15 \right] \right]$. The value that closes $i_1$ is a trapezoid in the form $\left[ \alpha, \beta, \gamma, \delta \right]$.

@@

\begin{eqnarray}
\label{eq:pvp-close-current}
\text{Close-current} \left(r, t \right) =\\
\begin{cases}
\nonumber
r - t_{CUR} \cup \lbrace t_{UP}\rbrace \cup \lbrace t \rbrace
& \mbox{ if } \mbox{Current } \left(r, t[A_K] \right) \neq \emptyset \\
% R - \left(A, I \right) \cup \left \lbrace \left(A, \mbox{ CloseR } \left(I, J\right) \right)\right \rbrace \cup \left \lbrace\left(A, J\right) \right \rbrace  \\
% \nonumber
% \mbox{ if  Current } \left(R, A \right) \neq \emptyset\\ %\wedge J > \mbox{ CloseR } \left(I, J \right)   \\
\nonumber r , \text{ otherwise}
\end{cases}
\end{eqnarray}
\end{definition}

@@

\item The entity is in the relation. Depending on the value of the time interval $i$, there are three possibilities:
	\begin{enumerate}
	\item Insert $t$ in the instance $r$ of the relation $R_{FTG}$. If the time interval $i$ does not overlap any other valid-time interval in the relation $R_{FTG}$. Note that here, the result of the overlaps operator is in the unit interval.
For example, the document with ID=3 is still valid. The insert sentence is the following.

@@

	\item Reject the insertion, if the time interval $i$ does overlap with a degree of 1 any existing valid-time interval for the entity in the relation. For example, consider now that the document manager wants to introduce a past validity period for the document with ID=3. The validity starting date for the document was around 6/4/12/1012 and the ending date was around 25/4/1012. As this interval does overlaps other time intervals with a degree of 1, it is not possible that the document was valid during two different periods of time. Therefore, the insertion is rejected. The insert sentence is:

@@

\begin{align}
\label{eq:insert-fuzzy-temporal}
\INS \left(r, t\right) =\\
\begin{cases}
\nonumber
r & \mbox{ if }  \exists t_k \in V(t), \\
\nonumber &\left(t[S,E] \mbox{ overlaps } t_k[S,E]=1\right)\\
\nonumber
r \cup \lbrace t \rbrace
& \mbox{ if }  V(t)=\emptyset \mbox{ or }\\
\nonumber & \forall t_k \in V(t), \\
\nonumber & \left(t[S,E] \mbox{ overlaps } t_k[S,E]\right) < 1\\
% 
% R \cup \left(A, I \right) & \mbox{ if }  (A, I ) \not \in R  \mbox{ or } \\
%  &   \left(A, J\right) \in R \mbox{ and } \lnot \left(I \mbox{ Overlaps } J\right) \\
\mbox{modify(r, t)} & \mbox{ otherwise }    %\mbox{ if } \exists J \left(A,J\right) \in R \wedge \mbox{Current(A,J)} \\ 
%R, &   
\end{cases} 	
\end{align}

\subsubsection{\label{subsubsec:delete-fuzzy-temporal}Delete}
The delete operation logically removes an entity which is valid in the instance $r$ of the relation $R_{FTG}$.

@@

\begin{align}
\label{eq:delete-fuzzy-temporal}
\DEL \left(r, t\right) =
\nonumber
r - V(t)
\end{align}
